---
title: "Superhero EDA - In Progress"
output:
  github_document:
    toc: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Introduction
It feels as though superhero movies have taken over the box office. It's hard to name a popular actor who hasn't appeared in at least one film where people can fly or move things with their minds. The goal of this project is to dig deeper into the growth of superhero movies and ultimately determine what percentage of popular actors have been in such films. This analysis is extremely subjective, but it does give me a great opportunity to utilize the `purrr` package to automate API calls.

All of the code used to conduct this analysis will be available in this repository. More in-depth explanations of the code used will be available in various posts on my website.[https://mhdemoblog.netlify.app/](https://mhdemoblog.netlify.app/)

## tl;dr
Superhero movies have become more expensive, profitable, and abundant over the last 15 years. Nearly 1 in every 3 active famous actors has been in a superhero movie. Stan Lee was a very busy man.

## TMDb API 
All of the data used in this analysis, will be queried from The Movie Database. This site offers an API that is free to access with no rate limiting. More information on the TMDb API can be found on their fact page.

[TBDb API Documentation](https://www.themoviedb.org/documentation/api)

## Packages
```{r, message=FALSE}
library(tidyverse)
library(glue)
library(jsonlite)
library(janitor)
library(lubridate)
library(scales)
library(patchwork)
library(ggfittext)
library(xfun)

```


## The Queries Pt 1
Before I can conduct my analysis I will need to query the data using the API. So what's the best way to start?

### Searching for Keywords
First I need to find all of the keywords that contain the word "superhero" and save the IDS for those keywords.
```{r}
keyword_search <- "superhero"

fromJSON(glue("https://api.themoviedb.org/3/search/keyword?\\
                api_key={Sys.getenv('THE_MOVIE_DB_KEY')}\\
                &query={keyword_search}")) %>%
  pluck(2) %>% 
  head()

super_hero_key_ids <- fromJSON(glue("https://api.themoviedb.org/3/search/keyword?\\
                api_key={Sys.getenv('THE_MOVIE_DB_KEY')}\\
                &query={keyword_search}")) %>%
  pluck(2) %>% 
  pull(2)
```
### Superhero Movie Ids
Using the keyword IDS I can search for any movies that contain those keywords. I just need to determine how many pages of results there are, create a function that queries the individual pages, and use the `map` function to iterate over those pages.
```{r}
# without_genres=16 - removes movies classified as animation
# with_release_type=3 - filters for theatrical releases
(page_count <- fromJSON(glue("https://api.themoviedb.org/3/discover/movie?\\
                       api_key={Sys.getenv('THE_MOVIE_DB_KEY')}\\
                       &certification_country=US&language=en-US&\\
                       without_genres=16&\\
                       with_release_type=3\\
                       &region=US\\
                       &with_keywords=\\
                       {glue_collapse(super_hero_key_ids, sep = '|')}\\
                       &primary_release_date.gte=1970-01-01")) %>%
  pluck(3))
```

```{r}
super_discover_query <- function(page_num) {
  fromJSON(glue("https://api.themoviedb.org/3/discover/movie?\\
                       api_key={Sys.getenv('THE_MOVIE_DB_KEY')}\\
                       &certification_country=US&language=en-US&\\
                       without_genres=16&\\
                       with_release_type=3\\
                       &region=US\\
                       &with_keywords=\\
                       {glue_collapse(super_hero_key_ids, sep = '|')}\\
                       &primary_release_date.gte=1970-01-01&page={page_num}")) %>%
    pluck(4)
}

super_discover_query(1) %>%
  glimpse()
```

```{r, include=FALSE}
# Stores the first two movies from the first page of the query so
# that the markdown text is up to date
sh_movies_p1 <- super_discover_query(1) %>% 
  slice(1:2) %>% 
  pull(title)
```

The function appears to work as the first two movies visible from the `title` column are <i>`r sh_movies_p1[1]`</i> and <i>`r sh_movies_p1[2]`</i>. Next up is the iteration.

```{r}
hero_movie_id <- map_df(1:page_count,
                        ~super_discover_query(.x)) %>%
  pull(id)

hero_movie_id %>% 
  glimpse()
```

The results from the pages were combined into a single data frame. I then used the `pull` function again to extract a vector containing the movie ids. It looks as though there have been `r length(hero_movie_id)` movies released in American theaters since Jan 1, 1970 that have been classified as superhero films.

### Movie Details
I could use the rest of the data found in the discover API call, but I want to know more. Have the budgets for superhero movies changed since 1970? Have they become more profitable? Additional information on these movies' cost and performance can be found using a slightly different API call. Querying the data requires a similar process as collecting all of the movie IDS.

```{r}
super_detail_query <-  function(movie_id) {

  fromJSON(glue("https://api.themoviedb.org/3/movie/{movie_id}?\\
                api_key={Sys.getenv('THE_MOVIE_DB_KEY')}&language=en-US"))
}

super_detail_query(hero_movie_id[1]) %>%
  glimpse()
```

Wow! This list contains a combination of lists and data frames. I can see some fields that I'm interested in, such as `budget` and `vote_average`. I just need to map over the movie IDS and transform these list elements into a single data frame.

```{r super-movies}
super_movies <- xfun::cache_rds({
  map(hero_movie_id,
       ~super_detail_query(.x)) %>%
    map_df(~as_tibble(t(.x))) %>%
    select(budget, imdb_id, original_title, title,
           release_date, revenue, runtime, status, vote_average,
           vote_count, id) %>%
    unnest(cols = everything()) %>%
    rename(movie_id = id) %>% 
    mutate(release_date = date(release_date))
})
```

```{r}
super_movies %>%
  glimpse()
```

This looks much easier to interpret. All of the information that I want is in a nice data frame and ready for some exploratory data analysis.

## Super Analysis
How many superhero movies have been released per year since 1970? Have these movies become more profitable over time? I'll attempt to answer these questions with some simple plots. I've also chosen to adjust the financial figures for inflation.

```{r}
super_movies %>%
  mutate(release_year = year(release_date)) %>%
  filter(release_year < year(Sys.Date())) %>%
  count(release_year) %>%
  ggplot(aes(release_year, n)) +
  geom_line() + geom_point() +
  scale_y_continuous(breaks = breaks_width(1)) +
  scale_x_continuous(breaks = breaks_width(5)) +
  theme(panel.grid.minor.y = element_blank()) +
  labs(x = "Release Year", title = "Superhero Movies Released by Year")
```

There has definitely been an upward trend in the number of superhero movies released per year. More of these types of films start to pop up around the mid 2000s. But were these newer superhero films more successful than their older counterparts?

```{r, message=FALSE}
cpi_data <- read_csv("Data/cpi_data.csv")
curr_cpi <- cpi_data %>% 
  tail(1) %>% 
  pull(avg_cpi)
cpi_data <- cpi_data %>% 
  mutate(inf_rate = curr_cpi / avg_cpi) %>% 
  select(-avg_cpi)

super_movies_adj <- super_movies %>% 
  mutate(release_year = year(release_date)) %>%
  left_join(cpi_data, by = c("release_year" = "year")) %>% 
  mutate(across(.cols = c(budget, revenue),
                .fns = ~ inf_rate * .x,
                .names = "{col}_inf_adj"),
         profit = revenue_inf_adj - budget_inf_adj,
         roi = profit / budget_inf_adj)

super_summ <- super_movies_adj %>%
  filter(release_date < Sys.Date(), budget_inf_adj > 100000) %>%
  mutate(profit = revenue_inf_adj - budget_inf_adj,
         roi = profit / budget_inf_adj) %>%
  group_by(release_year) %>%
  summarize(across(.cols = c(budget_inf_adj, revenue_inf_adj, profit, roi),
                   .fns = median,
                   .names = "med_{col}"),
            .groups = "drop") %>%
  pivot_longer(cols = contains("med"), names_to = "metric")

super_summ %>%
  filter(metric != "med_roi", release_year < year(Sys.Date())) %>%
  ggplot(aes(release_year, value, col = metric)) +
  geom_line() + geom_point() +
  scale_y_continuous(labels = dollar, breaks = breaks_width(250000000)) +
  scale_x_continuous(breaks = breaks_width(5)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(x = "Release Year", col = "Metric", title = "Median Performance by Year")

super_summ %>%
  filter(metric == "med_roi", release_year < year(Sys.Date())) %>%
  ggplot(aes(release_year, value)) +
  geom_line() + geom_point() +
  scale_x_continuous(breaks = breaks_width(5)) +
  scale_y_continuous(labels = percent) +
  labs(x = "Release Year", y = "ROI", title = "Median ROI by Year")
```
There appear to be some early superhero films that performed great at the box office and offered investors some amazing returns. However, the highest returns seem to be held by movies released after 2012. There also appear to be less occurrences of annual losses for this genre after 2012. This could be a product of studios' improved understanding of how to produce and market these types of movies.

Before I move on to collecting the cast information, I'd like to know what the top performing movie of each decade was.

```{r}
super_movies_adj %>% 
  filter(budget > 100000, release_date < Sys.Date()) %>% 
  mutate(decade = glue("{release_year - (release_year %% 10)}'s")) %>% 
  group_by(decade) %>% 
  slice_max(order_by = profit,
            n = 1) %>% 
  ggplot(aes(decade, profit, label = glue("{title}\n{percent(roi)} ROI"))) + geom_col() +
  geom_bar_text(reflow = TRUE) +
  scale_y_continuous(labels = dollar) +
  labs(x = "Decade", y = "Profit")
```

This plot is pretty interesting. <i>Avengers: Endgame</i> is by far the greatest superhero success at the box office, but 1989's <i>Batman</i> offered investors a higher rate of return at a whopping 1,075 % ROI. The 2020s have only shown negative returns for superhero films at the time of this analysis.

## The Queries Pt 2
The movie data that I’ve collected so far is great, but it lacks any information on the cast for each film. I’ll have to use a slightly different set of API calls to collect all of the cast listing for the movies I’ve identified. 

```{r super-cc}
cast_query <- function(movie_id) {
  fromJSON(glue("https://api.themoviedb.org/3/movie/{movie_id}\\
                /credits?api_key={Sys.getenv('THE_MOVIE_DB_KEY')}"))
}

super_cc <- xfun::cache_rds({
  map(hero_movie_id, ~cast_query(.x))
})

super_cast <- super_cc %>% 
  map_df(~as_tibble(t(.x))) %>% 
  select(-crew) %>% 
  unnest(c(id, cast), names_repair = "unique") %>% 
  rename(movie_id = 1, actor_id = 6, actor = 7) %>% 
  select(movie_id:order)
```

Now that I have the cast data I can join it together with the rest of the movie data.

```{r}
super_data <- super_cast %>% 
  left_join(super_movies_adj,
            by = "movie_id") %>% 
  relocate(movie_id, title, budget_inf_adj, revenue_inf_adj, profit, character)
```


Before I move on to the last step, I'd like to answer a couple more questions. First up, what actor has been in the most superhero movies? This question should be pretty easy to answer for any MCU fan.

```{r}
super_data %>% 
  count(actor) %>% 
  arrange(desc(n)) %>% 
  head()

super_data %>% 
  filter(actor == "Stan Lee") %>% 
  slice_max(order_by = release_date, n = 1) %>% 
  select(title, release_date, actor, character)
```


Stan Lee has made a cameo in nearly every Marvel film with his last being <i>Avengers: Endgame</i>. His cameos were always a great touch in the films. The other actors below Mr. Lee in appearances makes sense except for one. Who is this Jimmy Star guy? If I dig a little deeper I might be able to find out.

```{r}
jimmy_credits <- pluck(fromJSON(glue("https://api.themoviedb.org/3\\
                               /person/1781358/movie_credits?\\
                               api_key={Sys.getenv('THE_MOVIE_DB_KEY')}&\\
                               language=en-US")), 1)

nrow(jimmy_credits)

jimmy_credits %>% 
  arrange(desc(vote_average)) %>% 
  select(title, release_date, character, vote_average) %>% 
  head()
```

Mr. Star appears to be a very active extra. He's had 12 appearances to date in superhero films and 86 across all genres! Above are the top five movies that he's appeared in by vote average.

## The Finale

The last thing that I need is collect a list of popular people as defined by TMDb.

```{r pop-actors}
pop_pages <- fromJSON(glue("https://api.themoviedb.org/3/person/popular\\
              ?api_key={Sys.getenv('THE_MOVIE_DB_KEY')}&language=en-US")) %>% 
  pluck(3)

# Function to query the data from the popular search
pop_query <- function(page_num) {

fromJSON(glue("https://api.themoviedb.org/3/person/popular\\
              ?api_key={Sys.getenv('THE_MOVIE_DB_KEY')}&language=en-US&page={page_num}")) %>% 
    pluck(4)
}

# Maps pop_query functions over the total number of a pages
pop_actors <- xfun::cache_rds({
  map_df(.x = c(1:pop_pages),
         ~pop_query(.x))
})

pop_actors %>% 
  select(popularity:name) %>% 
  slice_sample(n = 10)
```

This list is great but it contains some popular people that I don't want. I only want to focus current, popular, movie actors. I decided to filter for actors who have been in a American movie since 2005 and who are in the 75th percentile based upon the popularity metric.

```{r, message=FALSE}
top_pop_perf <- pop_actors %>%
  # Unnest the known_for data frames for each actor
  unnest(known_for, names_repair = "unique") %>% 
  rename(actor = 3, actor_id = 4,
         movie_id = 11) %>% 
  # Removes missing release date rows
  filter(release_date > 0) %>%  
  mutate(release_date = date(release_date)) %>% 
  group_by(actor) %>% 
  # Filters for the oldest known_for item for each actor
  slice_min(order_by = release_date, n = 1) %>% 
  ungroup() %>%
  filter(release_date > "2005-01-01", media_type == "movie", 
         known_for_department == "Acting", 
         original_language == "en") %>%
  # Creates a quantile value for each actor based upon their popularity
  mutate(q_rank = ntile(popularity, 4)) %>% 
  filter(q_rank == 4) %>% 
  select(actor, actor_id)

top_pop_perf %>% 
  nrow()

top_pop_perf %>% 
  slice_sample(n = 10)
```
After filtering I've narrowed the pool down to the top 734 actors. A sample of this pool should show more recognizable names.

I can use the setdiff function to identify the popular actors who have never been in a superhero film.

```{r}
non_supers <- top_pop_perf %>% 
  dplyr::setdiff(super_data %>% 
                   distinct(actor, actor_id))

(super_perc <- percent(1 - nrow(non_supers) / nrow(top_pop_perf)))

non_supers %>% 
  slice_sample(n = 15)

top_pop_perf %>% 
  slice_sample(n = 3)

```


After a little work I have my very rough estimate of how many popular actors have been in a superhero film, `r super_perc`. So roughly 1 out of every 3 popular actors is a pretty big proportion. Throughout this process I noticed some exceptions that seemed to crop up. Eva Mendes is categorized as a non-super but she starred in 2007's <i>Ghost Rider</i>, which isn't designated as a superhero film. Subjective factors like this have influenced every facet of this analysis.
